---
title: "R Profiler"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Why is My Code So Slow?

* Profiling is a systematic way to examine how much time is spent in different parts of a program
* Useful when trying to optimize your code
* Often code runs fine once, but what if you have to put it in a loop for 1,000 iterations? Is it still fast enough?
* Profilling is better than guessing

## On Optimizing Your Code

* Getting biggest impact on speeding up code depends on knowing where the code spends spends most of its time
* This cannot be done with performance analysis or profiling

> We should forget small efficiencies, say about 97% of the time:
> premature optimization is the root of all evil.
>
> --Donald Knuth

In other words, focus on writing code that works and is readable before optimizing.

## General Principles of Optimization

* Design first then optimize
* Measure (collect data), don't guess

## system.time()

* `system.time()` takes an arbitray R expression as input (can be wrapped in `{}`) and returns the amount of time taken to evaluate the expression
* Computes the time (in seconds) needed to execute an expression
  - If there's an error, gives time until error occurred
* Returns an object of class `proc_time`
  - **user time**: time charged to the CPU(s) for this expression
  - **elapsed time**: "wall clock" time
* Usually, the user time and elapsed time are close for straight computing tasks
* Elapsed time may be *greater than* user time if the CPU spends a lot of time waiting around
* Elapsed time may be *smaller than* user time if your machine has multiple cores (and is capable of using them)
  - Multi-threaded BLAS libraries
  - Parallel processing via the `parallel` package

## system.time() Examples

```{r}
# elapsed time > user time
system.time(readLines("http://www.jhsph.edu"))

# elapsed time < user time
hilbert <- function(n) {
  i <- 1:n
  1 / outer(i - 1, i, "+")
}
x <- hilbert(1000)
system.time(svd(x))
```

## Timing Longer Expression

```{r}
system.time({
  n <- 1000
  r <- numeric(n)
  for (i in 1:n) {
    x <- rnorm(n)
    r[i] <- mean(x)
  }
})
```

## Beyond system.time()

`system.time()` is great when you know where the problem is occurring, or you are working with a small block of code, but what if you don't know where to start?

## The R Profiler

* The `Rprof()` function starts the profiler in R
   - R must be compiled with profiler support (generally this is true)
* The `summaryRprof()` function summarizes the output of `Rprof()` (it's unreadable otherwise)
* **DO NOT** use `system.time()` and `Rprof()` together!
* `Rprof()` keeps track of the function call stack at regularly sampled intervals and tabulates how much time is spent in each function
* Default sampling interval is 0.02 seconds
  - Therefore if your code runs very quickly the profiler will not be useful
  
## Using summaryRprof()

* The `summaryRprof()` function tabulates the R profiler output and calculates how much time is spent in which function
* There are two methods for normalizing the data
* `"by.total"` divides the time spent in each function by the total run time
* `"by.self"` does the same but first subtracts out time spent in functions above in the call stack
