---
title: "Loop Functions - apply"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## apply

`apply` is used to evaluate a function over the margins of an array.

* It is most often used to apply a function to the rows or columns of a matrix
* It can be used with general arrays, e.g. taking the average of an array of matrices
* It is not really faster than writing a loop, but it works in one line

```{r}
str(apply)
```

* `x` is an array
* `MARGIN` is an integer vector indicating which margins should be "retained"
* `FUN` is a function to be applied
* `...` is for other arguments to be passed to `FUN`

```{r}
x <- matrix(rnorm(200), 20, 10)
apply(x, 2, mean) # gets the mean of each column
apply(x, 1, sum) # gets the sum of each row
```

## col/row sums and means

For sums and means of matrix dimensions, we have some shortcuts.

* `rowSums` = `apply(x, 1, sum)`
* `rowMeans` = `apply(x, 1, mean)`
* `colSums` = `apply(x, 2, sum)`
* `colMeans` = `apply(x, 2, mean)`

These shortcuts are *much* faster than `apply` when using large matrices.

## Working in More Than 2 Dimensions

Let's say we want the average matrix of an array of matrices. Below we have an array of 10 $2\times2$ matrices (total dimensions $2\times2\times10$), we can pass a vector to the `MARGIN` argument specifying to retain the first two dimensions and collapsing the third.

```{r}
a <- array(rnorm(2 * 2 * 10), c(2, 2, 10))
dim(a) # array of 10 2x2 matrices

apply(a, c(1, 2), mean)

# equivlent call using rowMeans
rowMeans(a, dims = 2)
```

