---
title: "Loop Functions - lapply"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Looping on the Command Line

Writing `for`, `while` loops is useful when programming but not very convenient/concise when working interactively on on the command line. There are some functions which implement looping to make life easier.

* `lapply`: loop over a list and evaluate a function on each element
* `sapply`: same as `lapply` but try to simplify the result
* `apply`: apply a function over the margins of an array
* `tapply`: apply a function over subsets of a vector
* `mapply`: multivariate version of `lapply`

An auxiliary function `split` is also useful, particularly in conjunction with `lapply`.

## lapply

`lapply` takes three arguments: 

1. a list `x`
2. a function (or the name of a function) `FUN`
3. other arguments via `...`

Note that `lapply` will coerce `x` to a list using `as.list`.

```{r}
lapply
```

The actual looping is done internally in C code.

`lapply` always returns a list, regardless of class input.

```{r}
x <- list(a = 1:5, b = rnorm(10))
lapply(x, mean)
```

When using `lapply` you can also specify arguments of the function (`FUN`) that you are applying.

```{r}
x <- 1:4
lapply(x, runif, min = 0, max = 10)
```

## Anonymous Functions

`lapply` and related functions make heavy use of *anonymous* functions. Anonymous functions are just functions defined and called without being named. These are convenient when a function doesn't already exist to perform the operation needed and you have to write it yourself. 

```{r}
x <- list(a = matrix(1:4, 2, 2), b = matrix(1:6, 3, 2))
lapply(x, function(elt) elt[, 1]) ## anonymous function
```

## sapply

`sapply` will try to simplify the result of `lapply` if possible.

* If the result is a list where every element is length 1 then a vector is returned
* If the result is a list where every element is a vector of the same length (>1), a matrix is returned
* If no simplifications can be made then a list is returned

```{r}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20, 1), d = rnorm(100, 5))
y <- lapply(x, mean)
z <- sapply(x, mean)
class(y)
class(z)
```

