---
title: "Loop Functions - split"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## split

`split` is not a loop function, but is useful when combined with the `apply` family. This function takes a vector or other objects and splits it into groups determined by a factor or list of factors.

```{r}
str(split)
```

* `x` vector/list or data frame
* `f` factor (or coerced to one) or a list of factors
* `drop` indicates whether empty factors levels should be dropped

`split` always returns a list.

```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1))
f <- gl(3, 10)
split(x, f)
```

A common idiom is to use `split` followed by an `lapply`.

```{r}
split(x, f) |>
  lapply(mean)
```

## Splitting a Data Frame

```{r}
library(datasets)
head(airquality)
```

One thing we might want to do is calculate the means of each column, grouped by each month.

```{r}
split(airquality, airquality$Month) |>
  sapply(function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")], 
                              na.rm = TRUE))

```

## Splitting on More Than One Level

```{r}
x <- rnorm(10)
f1 <- gl(2, 5)
f2 <- gl(5, 2)
interaction(f1, f2)
```

Interactions can create empty levels.

```{r}
str(split(x, list(f1, f2)))
```

Specifying `drop = TRUE` will remove these empty levels.

```{r}
str(split(x, list(f1, f2), drop = TRUE))
```

