---
title: "Dates and Times in R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R has developed a special representation of dates and times.

* Dates are represented by the `Date` class
* Times are represented by the `POSIXct` or the `POSIXlt` class
* Dates are stored internally as the number of days since 1970-01-01
* Times are stored internally as the number of seconds since 1970-01-01

## Dates in R

Dates are represented by the `Date` class and can be coerced from a character string using the `as.Date()` function.

```{r}
x <- as.Date("1970-01-01")
x
unclass(x)
unclass(as.Date("1969-01-01"))
```

Note that when printing a date in R, it appears as a character string. It is **NOT** a character string, there is just a special print function for dates.

## Times in R

Times are represented by the `POSIXct` or the `POSIXlt` class.

* `POSIXct` is just a very large integer under the hood; it is a useful class when you want to store times in something like a data frame
* `POSIXlt` is a list underneath and it stores a bunch of other useful information like the day of the week, day of the year, etc.

There are a number of generic functions that work on dates and times.

* `weekdays`: give the day of the week
* `months`: give the month name
* `quarters`: give the quarter name ("Q1", "Q2", "Q3", or "Q4")

Times can be coerced from a character string using the `as.POSIXlt` or `as.POSIXct` function.

```{r}
x <- Sys.time()
p <- as.POSIXlt(x)
names(unclass(p))
p$sec
```

You can also use the `POSIXct` format.

Finally, there is the `strptime` function in case your dates are written in a different format.

```{r}
datestring <- c("January 10, 2012 10:40", "December 9, 2011 9:10")
x <- strptime(datestring, "%B %d, %Y %H:%M")
x
class(x)
```
For formatting help use `?strptime`.

## Operations on Dates and Times

You can use mathematical operations on dates and times. Addition, subtraction, and logical comparisons are all allowed.

```{r}
x <- as.Date("2012-01-01")
y <- strptime("9 Jan 2011 11:34:21", "%d %b %Y %H:%M:%S")
# x-y throws an error for incompatible methods
x <- as.POSIXlt(x)
x-y
```

A very nice benefit of these operations is they track things like leap years, leap seconds, daylight savings, and time zones.
