---
title: "Textual Data Formats"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Textual Formats

* `dump` and `dput` are useful because the resulting textual format is edit-able, and in the case of corrpution, potentially recoverable
* Unlike writing out a table or csv file, `dump` and `dput` preserve the *metadata* (sacrificing some readability), so that another user doesn't have to specify it all over again
* Textual formats can work much better with version control programs like subversion or git which can only track changes meaningfully in text files
* Textual formats can be longer-lived; if there is corruption somewhere in the file, it can be easier to fix the problem
* Textual formats adhere to the "Unix philosophy"
* Downside: the format is not very space-efficient

## dput-ing R Objects

Another way to pass data around is deparsing the R object with `dput` and reading it back in using dget.

```{r}
y <- data.frame(a = 1, b = "a")
dput(y)
dput(y, file = "y.R")
new.y <- dget("y.R")
new.y
```

## dump-ing R Objects

Multiple objects can be deparsed using the `dump` function and read back in using `source`.

```{r}
x <- "foo"
y <- data.frame(a = 1, b = "a")
dump(c("x", "y"), file = "data.R")
rm(x, y)
source("data.R")
y
x
```


