---
title: "Scoping Rules - Optimization"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Application: Optimization

Why is any of this information useful?

* Optimization routines in R like `optim`, `nlm`, and `optimize` require you to pass a function whose argument is a vector of parameters (e.g. a log-likelihood).
* However, an object function might depend on a host of other things besides its parameters (like data).
* When writing software which does optimization, it may be desirable to allow the user to hold certain parameters fixed.

## Maximizing a Normal Likelihood

Write a "constructor" function.

```{r}
make.NegLogLik <- function(data, fixed = c(FALSE, FALSE)) {
  params <- fixed
  function(p) {
    params[!fixed] <- p
    mu <- params[1]
    sigma <- params[2]
    a <- -0.5*length(data)*log(2*pi*sigma^2)
    b <- -0.5*sum((data-mu)^2) / (sigma^2)
    -(a + b)
  }
}
```

Note: optimization functions in R *minimize* functions, so you need to use the negative log-likelihood.

```{r}
set.seed(1); normals <- rnorm(100, 1, 2)
nLL <- make.NegLogLik(normals)
nLL
ls(environment(nLL))
```

## Estimating Parameters

```{r}
optim(c(mu = 0, sigma = 1), nLL)$par
```

Fixing $\sigma = 2$

```{r}
nLL <- make.NegLogLik(normals, c(FALSE, 2))
optimize(nLL, c(-1, 3))$minimum
```

Fixing $\mu = 1$

```{r}
nLL <- make.NegLogLik(normals, c(1, FALSE))
optimize(nLL, c(1e-6, 10))$minimum
```

## Plotting the Likelihood

```{r}
nLL <- make.NegLogLik(normals, c(1, FALSE))
x <- seq(1.7, 1.9, len = 100)
y <- sapply(x, nLL)
plot(x, exp(-(y-min(y))), type = "l")
```

```{r}
nLL <- make.NegLogLik(normals, c(FALSE, 2))
x <- seq(0.5, 1.5, len = 100)
y <- sapply(x, nLL)
plot(x, exp(-(y-min(y))), type = "l")
```


## Lexical Scoping Summary

* Objective functions can be "built" which contains all the necessary data for evaluating thee function.
* No need to carry around long argument lists - useful for interactive and exploratory work.
* Code can be simplified and cleaned up
