---
title: "Scoping Rules - Symbol Binding"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A Diversion on the Binding Values to Symbol

How does R know which value to assign to which symbol?

```{r}
lm <- function(x) {x * x}
lm
```

Note that there already exists a function in R called `lm`. How does R know what value to assign to the symbol `lm`? Why doesn't it give it the value of `lm` that is in the *stats* package?

When R tries to bind a value to a symbol, it searches through a series of `environments` to find the appropriate value. When you are working on the command line and need to retrieve the value of an R object, the order is roughly

1. Search the global environment for a symbol name matching the one requested.
2. Search the namespaces of each of the packages on the search list. These are packages that have been loaded.

The search list can be found by using the `search` function.

```{r}
search()
```

## Binding Values to Symbol

* The *global environment* or the user's workspace is always the first element of the search list and the *base* package is always the last.
* The order of the packages on the search list matters!
* User's can configure which packages get loaded on startup so you cannot assume that there will be a set list of packages available.
* When a user loads a package with `library` the namespace of that package gets put in position 2 of the search list (by default) and everything else gets shifted down the list.
* Note that R has spearate namespaces for functions and non-functions so it's possible to have an object named `c` and a function named `c`.

## Scoping Rules

The scoping rules for R are the main feature that makes it different from the original S language.

* The scoping rules determine how a value is associated with a free variable in a function
* R uses *lexical scoping* or *static scoping*. A common alternative is *dynamic scoping*.
* Related to the scoping rules is how R uses the search *list* to bind a value to a symbol
* Lexical scoping turns out to be particularly useful for simplifying statistical computations

## Lexical Scoping

Consider the following:

```{r eval = FALSE}
f <- function(x, y) {
  x^2 + y / z
}
```

This function has 2 formal arguments `x` and `y`. In the body of the function there is another symbol `z`. In this case `z` is called a *free variable*. The scoping rules of a language determine how values are assigned to free variables. Free variables are not formal arguments and are not local variables (assigned inside the function body).

Lexical scoping in R means that 

> the values of free variables are searched for in the environment in which the function was defined.

What is an environment?

* An *environment* is a collection of (symbol, value) pairs, i.e. `x` is a symbol and `3.14` might be its value.
* Every environment has a parent environment; it is possible for an environment to have multiple "children".
* The only environment without a parent is the empty environment.
* A function + an environment = a *closure* or *function closure*.

Searching for a value for a free variable:

* If the value of a symbol is not found in the environment in which a function was defined then the search is continued in the *parent environment*.
* The search continues down the sequence of parent environments until we hit the *top-level environment*; this is usually the global environment (workspace) or the namespace of a package.
* After the top-level environment, the search continues down the search list until we hit the *empty environment*. If a value for a given symbol cannot be found once the empty environment is arrived at then an error is thrown. 

Why does this matter?

* Typically, a function is defined in the global environment, so that the values of the free variables are just found in the user's workspace.
* This behavior is logical for most people and is usually the "right thing" to do.
* However, in R you can have functions defined *inside other functions*:
  - Languages like C do not allow this
* Now things get interesting - in this case the environment in which a function is defined is the body of another function!

## Example

```{r}
make.power <- function(n) {
  pow <- function(x) {
    x^n
  }
  pow
}
```

This function returns another function as its value.

```{r}
cube <- make.power(3)
square <- make.power(2)
cube(3)
square(3)
```

## Exploring a Function Closure

What's in a function's environment?
```{r}
ls(environment(cube))
get("n", environment(cube))

ls(environment(square))
get("n", environment(square))
```

## Lexical vs. Dynamic Scoping

```{r eval = FALSE}
y <- 10

f <- function(x) {
  y <- 2
  y^2 + g(x)
}

g <- function(x) {
  x*y
}
```

What is the value of `f(3)`?

* With lexical scoping the value of `y` is the function `g` is looked up in the environment in which the function was defined, in this case the global environment, so the value of `y` is 10.
* With dynamic scoping, the value of `y` is looked up in the environment from which the function was *called* (sometimes referred to as the *calling environment*).
  - In R the calling environment is known as the *parent frame*
* So the value of `y` would be 2.

When a function is *defined* in the global environment and *called* from the global environment then the defining and calling environment are the same. 

```{r eval = FALSE}
g <- function(x) {
  a <- 3
  x + a + y
}
g(2) # throws an error
y <- 3
g(2) # outputs 8
```

## Other Languages

Other languages that support lexical scoping:

* Scheme
* Python
* Perl
* Common lisp

## Consequences of Lexical Scoping

* In R, all objects must be stored in memory.
* All functions must carry a pointer to their respective defining environments, which could be anywhere.
* In S-PLUS, free variables are always looked up in the global workspace, so everything can be stored on the disk because the "defining environment" of all functions is the same.