---
title: "Scoping Rules - Symbol Binding"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A Diversion on the Binding Values to Symbol

How does R know which value to assign to which symbol?

```{r}
lm <- function(x) {x * x}
lm
```

Note that there already exists a function in R called `lm`. How does R know what value to assign to the symbol `lm`? Why doesn't it give it the value of `lm` that is in the *stats* package?

When R tries to bind a value to a symbol, it searches through a series of `environments` to find the appropriate value. When you are working on the command line and need to retrieve the value of an R object, the order is roughly

1. Search the global environment for a symbol name matching the one requested.
2. Search the namespaces of each of the packages on the search list. These are packages that have been loaded.

The search list can be found by using the `search` function.

```{r}
search()
```

## Binding Values to Symbol

* The *global environment* or the user's workspace is always the first element of the search list and the *base* package is always the last.
* The order of the packages on the search list matters!
* User's can configure which packages get loaded on startup so you cannot assume that there will be a set list of packages available.
* When a user loads a package with `library` the namespace of that package gets put in position 2 of the search list (by default) and everything else gets shifted down the list.
* Note that R has spearate namespaces for functions and non-functions so it's possible to have an object named `c` and a function named `c`.

## Scoping Rules

The scoping rules for R are the main feature that makes it different from the original S language.

* The scoping rules determine how a value is associated with a free variable in a function
* R uses *lexical scoping* or *static scoping*. A common alternative is *dynamic scoping*.
* Related to the scoping rules is how R uses the search *list* to bind a value to a symbol
* Lexical scoping turns out to be particularly useful for simplifying statistical computations

## Lexical Scoping

Consider the following:

```{r eval = FALSE}
f <- function(x, y) {
  x^2 + y / z
}
```

This function has 2 formal arguments `x` and `y`. In the body of the function there is another symbol `z`. In this case `z` is called a *free variable*. The scoping rules of a language determine how values are assigned to free variables. Free variables are not formal arguments and are not local variables (assigned inside the function body).

Lexical scoping in R means that 

> the values of free variables are searched for in the environment in which the function was defined.

What is an environment?

* An *environment* is a collection of (symbol, value) pairs, i.e. `x` is a symbol and `3.14` might be its value.
* Every environment has a parent environment; it is possible for an environment to have multiple "children".
* The only environment without a parent is the empty environment.
* A function + an environment = a *closure* or *function closure*.

Searching for a value for a free variable:

* If the value of a symbol is not found in the environment in which a function was defined then the search is continued in the *parent environment*.
* The search continues down the sequence of parent environments until we hit the *top-level environment*; this is usually the global environment (workspace) or the namespace of a package.
* After the top-level environment, the search continues down the search list until we hit the *empty environment*. If a value for a given symbol cannot be found once the empty environment is arrived at then an error is thrown. 