---
title: "Data Types - Factors"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Factors

Factors are used to represent categorical data. Factors can be unordered or ordered. One can think of a factor as an integer vector where each integer is *labeled*.

* Factors are treated specially by modelling fuctions like `lm()` and `glm()`
* Using factors with labels is *better* than using integers because factors are self-describing; having a variable that has values "Male" and "Female" is better than a variable that has values 1 and 2.

```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"))
print(x)
print(table(x))
print(unclass(x))
```

The order of the levels can be set using the `levels` argument to `factor()`. This can be important in linear modeling because the first level is used as the baseline level. **NOTE**: The default choice for levels is alphabetical order.

```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"),
            levels = c("yes", "no"))

x
```

