---
title: "Data Types - Vectors and Lists"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Creating Vectors

The `c()` function can be used to create vectors of objects.

```{r}
x <- c(0.5, 0.6)      ## numeric
x <- c(TRUE, FALSE)   ## logical
x <- c(T, F)          ## also logical
x <- c("a", "b", "c") ## character
x <- 9:29             ## integer
x <- c(1+0i, 2+4i)    ## complex
```

Using the `vector()` function

```{r}
x <- vector("numeric", length = 10)
x
```


## Mixing Ojbects

What happens below?

```{r}
y <- c(1.7, "a")  ## character
y <- c(TRUE, 2)   ## numeric
y <- c("a", TRUE) ## character
```

R won't error, but will instead *coerce* every element in the vector to the same class.

## Explicit Coercion

Objects can be coerced explicitly using the `as.*` functions, if available.

```{r}
x <- 0:6
print(class(x))
print(as.numeric(x))
print(as.logical(x))
print(as.character(x))

```

Nonsensical coercion results in `NA`s.

```{r}
x <- c("a", "b", "c")
print(as.numeric(x))
print(as.logical(x))
print(as.complex(x))
```

## Lists

Lists are a special type of vector that can contain elements of different classes. Lists are a very important data type in R.

```{r}
x <- list(1, "a", TRUE, 1 + 4i)
x
```

